The Mass Distribution of Stellar-Mass Black Holes

### 2010-08-06 paper.tex
In-progress major edits, new plots Makefile, some new plots

\usepackage{hyperref}
\usepackage{amsmath}

\newcommand{\fixme}[1]{[FIXME: #1]}
\newcommand{\Msun}{M_\odot}
\newcommand{\Mmin}{M_{\textnormal{min}}}
\newcommand{\Mmax}{M_{\textnormal{max}}}
\newcommand{\Nbin}{N_{\textnormal{bin}}}

@@

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes based on the observed masses of the compact
  object in 17 low mass X-ray binaries.  Modeling the mass
  distribution both parametrically---as a power law, exponential,
  gaussian, or combination of two gaussians---and
  non-parametrically---as histograms with varying numbers of
  bins---we explore the distribution in the context of Markov Chain
  Monte Carlo calculations of model parameters.  We give confidence
  bounds on the shape of the mass distribution in the context of each
  model and compare the models with each other by calculating their
  Bayesian evidence.  We address the existence of a ``gap'' between
  the most massive neutron stars and least massive black holes,
  finding that the best model (the power law), the second-best model
  (a one-bin histogram, or flat distribution), and the third-best
  model (a two-bin histogram) have minimum black hole masses above
  5.00, 3.23, and 3.82 solar masses (95\% confidence), respectively.
  We therefore conclude that our sample of black hole masses provides
  strong evidence of a gap between the maximum neutron star mass and
  minimum black hole mass.
\end{abstract}

@@

The 17 low mass X-ray binary systems on which this study is based are
listed in Table \ref{tab:sources}.  In each of these systems,
spectroscopic measurements of the secondary star provide an orbital
period for the system and a half-amplitude for the secondary's
velocity curve.  These measurements can be combined into the mass
function,
\begin{equation}
  \label{eq:mass-function}
  f(M) = \frac{P K^3}{2\pi G} = \frac{M \sin^3 i}{\left( 1 + q \right)^2},
\end{equation}
where $P$ is the orbital period, $K$ is the secondary's velocity
semi-amplitude, $M$ is the black hole mass, $i$ is the inclination of
the system, and $q \equiv M_2 / M$ is the mass ratio of the system.

The mass function defines a lower limit on the mass: $f(M) < M$.  To
accurately determine the mass of the black hole, the inclination $i$
and mass ratio $q$ must be measured.  Ideally, this can be
accomplished by fitting ellipsoidal light curves from these systems
but even in the ideal case (see, e.g., \citet{Cantrell2010} on A0620)
this procedure can be complicated by possible contributions from an
accretion disk and hot spots in the disk.  For some systems (e.g.\ GS
1354 \citep{Casares2009}) strong variability prevents determination of
the inclination from the lightcurve at all; in these cases an upper
limit on the inclination often comes from the observed lack of
eclipses in the lightcurve.  In general, accurately determining $q$
and $i$ requires a careful system-by-system analysis.

For the purposes of this paper, we adopt the following simplified
approach to the estimation of the black hole mass from the observed
data.  When an observable is well-constrained, we assume that the true
value is normally distributed about the measured value with a standard
deviation equal to the quoted observational error.  This is the case
for the mass function in all the systems we use, and for many systems'
mass ratios and inclinations.  When a large range is quoted in the
literature for an observable, we take the true value to be distributed
uniformly (for the mass ratio) or isotropically (for the inclination)
within the quoted range.  Table \ref{tab:sources} gives the assumed
distribution for the observables in the 17 systems we use.  We do not
attempt to deal with the systematic biases in the observational
determination of $f$, $q$, and $i$ in any realistic way, instead
leaving this for future work.

\begin{table}
  \begin{center}
    \begin{tabular}{|l|c|c|c|l|}
      \hline
      Source & $f$ ($\Msun$) & $q$ & $i$ (degrees) & References \\
      \hline \hline
      GRS 1915 & $N(9.5, 3.0)$ & $N(0.0857, 0.0284)$ & $N(70, 2)$ &
      \citet{Greiner2001} \\
      XTE J1118 & $N(6.44, 0.08)$ & $N(0.0264, 0.004)$ & $N(68, 2)$ &
      \citet{Gelino2008} \\ & & & & \citet{Harlaftis2005} \\
      XTE J1650 & $N(2.73, 0.56)$ & $U(0, 0.5)$ & $I(50, 80)$ &
      \cite{Orosz2004} \\
      GRS 1009 & $N(3.17, 0.12)$ & $N(0.137, 0.015)$ & $I(37, 80)$ &
      \cite{Filippenko1999} \\
      A0620 & $N(2.76, 0.036)$ & $N(0.06, 0.004)$ & $N(50.98, 0.87)$ &
      \citet{Cantrell2010} \\ & & & & \citet{Neilsen2008} \\
      GRO J0422 & $N(1.13, 0.09)$ & $U(0.076, 0.31)$ & $N(45, 2)$ &
      \citet{Gelino2003} \\
      Nova Mus 1991 & $N(3.01, 0.15)$ & $N(0.128, 0.04)$ & $N(54,1.5)$
      & \cite{Gelino2001} \\
      GRO J1655 & $N(2.73,0.09)$ & $N(0.3663, 0.04025)$ & $N(70.2,
      1.9)$ & \citet{Greene2001} \\
      4U 1543 & $N(0.25, 0.01)$ & $U(0.25, 0.31)$ & $N(20.7,1.5)$ & 
      \citet{Orosz2003} \\
      XTE J1550 & $N(7.73,0.4)$ & $U(0,0.04)$ & $N(74.7, 3.8)$ &
      \citet{Orosz2010} \\
      V4641 Sgr & $N(3.13,0.13)$ & $U(0.42,0.45)$ & $N(75,2)$ &
      \citet{Orosz2003} \\
      GS 2023 & $N(6.08, 0.06)$ & $U(0.056,0.063)$ & $I(66, 70)$ &
      \citet{Charles2006} \\
      & & & & \citet{Khargharia2010} \\
      GS 1354 & $N(5.73, 0.29)$ & $N(0.12,0.04)$ & $I(50, 80)$ & 
      \citet{Casares2009} \\
      Nova Oph 77 & $N(4.86,0.13)$ & $U(0, 0.053)$ & $I(60, 80)$ &
      \citet{Charles2006} \\
      GS 2000 & $N(5.01, 0.12)$ & $U(0.035, 0.053)$ & $I(43, 74)$ &
      \citet{Charles2006} \\
      \hline \hline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% High Mass Follows
      Cyg X1 & $N(0.251, 0.007)$ & $N(2.778, 0.386)$ & $I(23, 38)$ &
      \citet{Gies2003} \\ 
      M33 X7 & $N(0.46, 0.08)$ & $N(4.47, 0.61)$ & $N(74.6, 1)$ &
      \citet{Orosz2007} \\
      NGC 300 X1 & $N(2.6, 0.3)$ & $U(1.05, 1.65)$ & $I(60, 75)$ &
      \citet{Crowther2010} \\
      LMC X1 & $N(0.148, 0.004)$ & $N(2.91, 0.49)$ & $N(36.38, 2.02)$
      & \citet{Orosz2009} \\
      IC 10 X1 & $N(7.64, 1.26)$ & $U(0.7, 1.7)$ & $I(75, 90)$ & 
      \citet{Prestwich2007} \\
       & & & & \citet{Silverman2008} \\
       \hline
    \end{tabular}
  \end{center}
  \caption{\label{tab:sources} The source parameters for the 17 low
    mass X-ray binaries used in this work.  $f$ is the mass function for
    the compact object, $q$ is the mass ratio $M_2/M$, and $i$ is the
    inclination of the system to the line of sight.  We indicate the
    distribution used for the true parameters when computing the
    probability distributions for the masses of these systems:
    $N(\mu,\sigma)$ implies a Gaussian with mean $\mu$ and standard
    deviation $\sigma$, $U(a,b)$ is a uniform distribution between $a$ and
    $b$, and $I(\alpha,\beta)$ is an isotropic distribution between the
    angles $\alpha$ and $\beta$.  \fixme{Orosz for XTE J1550 needs a
      real reference \citep{Orosz2010}.}} 
\end{table}

From these assumptions, we can generate probability distributions for
the true mass of the black hole given the observations and errors via
the Monte Carlo method: drawing samples of $f$, $q$, and $i$ from the
assumed distributions and computing the mass implied by Equation
\eqref{eq:mass-function} gives samples of $M$ from the distribution
induced by the relationship in Equation \eqref{eq:mass-function}.
Mass distributions generated in this way for the 17 systems used in
this work are shown in Figure \ref{fig:all-masses}.  These mass
distributions constitute the ``observational data'' we will use in
Section \ref{sec:models} to explore various underlying mass
distributions.

@@

  \caption{\label{fig:all-masses} The mass distributions implied by
    Equation \eqref{eq:mass-function} and the assumed distributions on
    observational parameters given in Table \ref{tab:sources}.  The
    significant asymmetry and long tails in many of these
    distributions are the result of the non-linear relationship
    (Equation \eqref{eq:mass-function}) between $M$, $f$, $q$, and
    $i$.}
\end{figure}

In Figure \ref{fig:masses} we show the combination of these individual
mass distributions into one single distribution.  (Note that this
combination is not a good estimator of the true mass distribution of
black hole masses---see, e.g., \citet{Mandel2010}.  We include it here
only to give a sense of the range of masses of black holes in our
sample; we will come to estimators for the mass distribution.)

\begin{figure}
  \plotone{plots/masses}

  \caption{\label{fig:masses} The mass distributions in Figure
    \ref{fig:all-masses} combined into a single distribution. Note
    that this combination is not a good estimator of the true mass
    distribution of black hole masses in the galaxy---see, e.g.,
    \citet{Mandel2010}.  We include it here only to give a sense of
    the range and general shape of the mass distribution of black
    holes in our sample.}
\end{figure}

@@

In this section we apply a Bayesian analysis to various models for the
underlying mass distribution from which the systems in Table
\ref{tab:sources} were drawn.  The end result will be the probability
distribution for the parameters of each model implied by the data from
Section \ref{sec:systems} in combination with our prior assumptions
about the probability distribution for the parameters.  Bayes' rule
relates these quantities.  For a model with parameters $\theta$ in the
presence of data $d$, Bayes' rule states
\begin{equation}
  \label{eq:Bayes-rule}
  p(\theta | d) = \frac{p(d | \theta) p(\theta)}{p(d)}.
\end{equation}
Here, $p(\theta|d)$, called the posterior probability distribution
function, is the probability distribution for the parameters $\theta$
implied by the data $d$; $p(d|\theta)$, called the likelihood, is the
probability of observing data $d$ given that the model parameters are
$\theta$; $p(\theta)$, called the prior, reflects our estimate of the
probability of the various model parameters in the absence of any
data; and $p(d)$, called the evidence, is an overall normalizing
constant ensuring that 
\begin{equation}
  \int d\theta\, p(\theta|d) = 1,
\end{equation}
whence
\begin{equation}
  \label{eq:evidence-def}
  p(d) = \int d\theta\, p(d|\theta) p(\theta).
\end{equation}

In our context, the data are the mass distributions given in Section
\ref{sec:systems}: $d = \{ p_i(m)| i = 1, 2, \ldots, 17 \}$.  We
assume that the measurements in Section \ref{sec:systems} are
independent, so the complete likelihood is given by a product of the
likelihoods for the individual measurements.  For a model with
parameters $\theta$ that predicts a mass distribution $p(m|\theta)$
for black holes, we have
\begin{equation}
  \label{eq:likelihood-def}
  p(d|\theta) = \prod_i \int dm\, p_i(m) p(m|\theta).
\end{equation}
That is, the likelihood of an observation is the average over the
individual mass distribution implied by the observation, $p_i(m)$, of
the probability for a black hole of that mass to exist according to
the model of the mass distribution, $p(m | \theta)$.  We approximate
the integrals as averages of $p(m|\theta)$ over the Monte Carlo mass
samples drawn from the distributions in Table \ref{tab:sources} (also
see Figure \ref{fig:all-masses}):
\begin{equation}
  p(d|\theta) \approx \prod_i \frac{1}{N_i} \sum_{j = 1}^{N_i} p(m_{ij} | \theta),
\end{equation}
where $m_{ij}$ is the $j$th sample (out of a total $N_i$) from the
$i$th individual mass distribution.

Our calculation of the likelihood of each observation does not include
any attempt to account for selection effects in the observations.  We
simply assume (almost certainly incorrectly) that any black hole drawn
from the underlying mass distribution is equally likely to be
observed, leaving the analysis of selection effects to future work.

For a mass distribution with several parameters, $p(\theta | d)$ lives
in a multi-dimensional space.  Exploring the entirety of this space
systematically rapidly becomes prohibitive as the number of parameters
increases.  A more efficient way to explore the distribution $p(\theta
| d)$ is to use a Markov Chain Monte Carlo (MCMC) method.  MCMC
methods produce a chain (sequence) of parameter samples, $\{ \theta_i,
i = 1, \ldots \}$, such that a particular parameter, $\theta$, appears
in the sequence with a frequency equal to its probability according to
$p(\theta|d)$.  In this way, regions of parameter space where
$p(\theta|d)$ is large are sampled densely while regions where
$p(\theta|d)$ is small are effectively ignored.  A Markov chain has
the property that the transition probability from one element to the
next, $p(\theta_i \to \theta_{i+1})$, depends only on the value of
$\theta_i$, not on any previous values in the chain.

One way to produce a sequence of MCMC samples is via the following
algorithm, first proposed by \fixme{Metropolis citation needed}:
\begin{enumerate}
  \item Begin with the current sample, $\theta_i$.
  \item Propose a new sample, $\theta_p$, by drawing randomly from an
    arbitrary ``jump proposal distribution'' with probability
    $Q(\theta_i \to \theta_p)$.
  \item Compute the ``acceptance'' probability,
    \begin{equation}
      \label{eq:paccept}
      p_{\textnormal{accept}} \equiv
      \frac{p(\theta_p|d)}{p(\theta_i|d)} \frac{Q(\theta_p \to
        \theta_i)}{Q(\theta_i \to \theta_p)}
    \end{equation}
  \item With probability $\min(1,p_{\textnormal{accept}})$ set
    $\theta_{i+1} = \theta_p$; otherwise set $\theta_{i+1} =
    \theta_i$.
\end{enumerate}
Note that this algorithm is more likely to accept a proposed jump when
it increases the posterior (the first term in Equation
\eqref{eq:paccept}) and when it is to a location in parameter space
from which it is easy to return (the second term in Equation
\eqref{eq:paccept}).  As $i \to \infty$ the samples $\theta_i$ are
distributed according to $p(\theta|d)$.  In practice the number of
samples required before the chain appropriately samples $p(\theta|d)$
depends strongly on the jump proposal distribution; proposal
distributions that often propose jumps toward or within regions of
large $p(\theta|d)$ can be very efficient, while poor proposal
distributions can require prohibitively large numbers of samples
before convergence.

Once we have a chain of samples from $p(\theta|d)$---a distribution of
distributions---the distribution for any quantity of interest can be
computed.  For example, Figure \ref{fig:dists} shows the median
values, and the 10\%, and 90\% quantiles for value of the BH mass
probability distribution at various masses in each of the models
discussed in the following subsections.  Such a figure can be
generated from the MCMC output by computing the value of the
parameterized BH mass distribution at various masses for each
parameter sample output by the MCMC, and then plotting the quantiles
at each mass.

\begin{figure}
  \begin{center}
    \plottwo{plots/dist-parameteric}{plots/dist-non-parameteric}
  \end{center}
  \caption{\label{fig:dists} The median values of the black hole mass
    distribution, $p(m|\theta)$, at various masses implied by the
    posterior $p(\theta|d)$ for the models discussed in Sections
    \ref{subsec:parameteric-models} and
    \ref{subsec:non-parameteric-models}.  Error bars span the 10\% to
    90\% range.  Note that these ``distributions of distributions''
    are not necessarily normalized, and need not be ``shaped'' like
    the underlying model distributions.}
\end{figure}

@@

We defer discussion of the appropriateness of the different models
given the data (i.e.\ the ``goodness of fit'') until Section
\ref{sec:model-selection}.

>>>>

\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}

\newcommand{\fixme}[1]{[FIXME: #1]}
\newcommand{\Msun}{M_\odot}
\newcommand{\Mmin}{M_{\textnormal{min}}}
\newcommand{\Mmax}{M_{\textnormal{max}}}
\newcommand{\Nbin}{N_{\textnormal{bin}}}
\newcommand{\vtheta}{\vec{\theta}}

@@

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems and five higher-mass, wind-fed X-ray binary
  systems.  Modeling the mass distribution both parametrically---as a
  power law, exponential, gaussian, combination of two gaussians, or
  log-normal distribution---and non-parametrically---as histograms
  with varying numbers of bins---we explore the distribution in the
  context of Markov Chain Monte Carlo calculations of model
  parameters.  We give confidence bounds on the shape of the mass
  distribution in the context of each model and compare the models
  with each other by calculating their Bayesian evidence.  The mass
  distribution of the low-mass systems is best fit by a power-law,
  while the distribution of the complete sample is best fit by the
  exponential model.  We address the existence of a ``gap'' between
  the most massive neutron stars and least massive black holes,
  finding that the best model (the power law) fitted to the low-mass
  systems only gives a minimum black-hole mass of 4.7 $\Msun$ (90\%
  confidence), while the best model (the exponential) fitted to all 20
  systems gives a minumum black-hole mass of 4.4 $\Msun$ (90\%
  confidence).  We conclude that our sample of black hole masses
  provides strong evidence of a gap between the maximum neutron star
  mass and minimum black hole mass.
\end{abstract}

@@

The 20 X-ray binary systems on which this study is based are listed in
Table \ref{tab:sources}.  We separate the systems into 15 low-mass
systems in which the central black hole appears to be fed by
roche-lobe overflow from the secondary, and 5 higher-mass systems in
which the black hole is fed via winds (these systems all have a
secondary that appears more massive than the black hole).  The low-
and higher-mass systems undoubtedly have different evolutionary
tracks, and therefore it is reasonable that they would have different
black-hole mass distributions.  We will first analyze the 15 low-mass
systems alone (Sections \ref{sec:models} and
\ref{sec:model-selection}), and then the complete sample of 20 systems
(Section \ref{sec:higher-mass}).

In each of these systems, spectroscopic measurements of the secondary
star provide an orbital period for the system and a half-amplitude for
the secondary's velocity curve.  These measurements can be combined
into the mass function,
\begin{equation}
  \label{eq:mass-function}
  f(M) = \frac{P K^3}{2\pi G} = \frac{M \sin^3 i}{\left( 1 + q \right)^2},
\end{equation}
where $P$ is the orbital period, $K$ is the secondary's velocity
semi-amplitude, $M$ is the black hole mass, $i$ is the inclination of
the system, and $q \equiv M_2 / M$ is the mass ratio of the system.

The mass function defines a lower limit on the mass: $f(M) < M$.  To
accurately determine the mass of the black hole, the inclination $i$
and mass ratio $q$ must be measured.  Ideally, this can be
accomplished by fitting ellipsoidal light curves and study of the
rotational broadening of spectral lines from the secondary, but even
in the ideal case (see, e.g., \citet{Cantrell2010} on A0620) this
procedure is complicated.  In particular, contributions from an
accretion disk and hot spots in the disk can significantly distort the
measured inclination and mass ratios.  For some systems (e.g.\ GS 1354
\citep{Casares2009}) strong variability prevents determination of the
inclination from the lightcurve at all; in these cases an upper limit
on the inclination often comes from the observed lack of eclipses in
the lightcurve.  In general, accurately determining $q$ and $i$
requires a careful system-by-system analysis.

For the purposes of this paper, we adopt the following simplified
approach to the estimation of the black hole mass from the observed
data.  When an observable is well-constrained, we assume that the true
value is normally distributed about the measured value with a standard
deviation equal to the quoted observational error.  This is the case
for the mass function in all the systems we use, and for many systems'
mass ratios and inclinations.  When a large range is quoted in the
literature for an observable, we take the true value to be distributed
uniformly (for the mass ratio) or isotropically (for the inclination)
within the quoted range.  Table \ref{tab:sources} gives the assumed
distribution for the observables in the 17 systems we use.  We do not
attempt to deal with the systematic biases in the observational
determination of $f$, $q$, and $i$ in any realistic way; we are
currently investigating more realistic treatments of the errors
(including observational biases that can shift the peak of the true
mass distribution away from the ``best-fit'' mass in the
observations).  This careful treatment will appear in future work.

\begin{table}
  \begin{center}
    \begin{tabular}{|l|c|c|c|l|}
      \hline
      Source & $f$ ($\Msun$) & $q$ & $i$ (degrees) & References \\
      \hline \hline
      GRS 1915 & $N(9.5, 3.0)$ & $N(0.0857, 0.0284)$ & $N(70, 2)$ &
      \citet{Greiner2001} \\
      XTE J1118 & $N(6.44, 0.08)$ & $N(0.0264, 0.004)$ & $N(68, 2)$ &
      \citet{Gelino2008} \\ & & & & \citet{Harlaftis2005} \\
      XTE J1650 & $N(2.73, 0.56)$ & $U(0, 0.5)$ & $I(50, 80)$ &
      \cite{Orosz2004} \\
      GRS 1009 & $N(3.17, 0.12)$ & $N(0.137, 0.015)$ & $I(37, 80)$ &
      \cite{Filippenko1999} \\
      A0620 & $N(2.76, 0.036)$ & $N(0.06, 0.004)$ & $N(50.98, 0.87)$ &
      \citet{Cantrell2010} \\ & & & & \citet{Neilsen2008} \\
      GRO J0422 & $N(1.13, 0.09)$ & $U(0.076, 0.31)$ & $N(45, 2)$ &
      \citet{Gelino2003} \\
      Nova Mus 1991 & $N(3.01, 0.15)$ & $N(0.128, 0.04)$ & $N(54,1.5)$
      & \cite{Gelino2001} \\
      GRO J1655 & $N(2.73,0.09)$ & $N(0.3663, 0.04025)$ & $N(70.2,
      1.9)$ & \citet{Greene2001} \\
      4U 1543 & $N(0.25, 0.01)$ & $U(0.25, 0.31)$ & $N(20.7,1.5)$ & 
      \citet{Orosz2003} \\
      XTE J1550 & $N(7.73,0.4)$ & $U(0,0.04)$ & $N(74.7, 3.8)$ &
      \citet{Orosz2010} \\
      V4641 Sgr & $N(3.13,0.13)$ & $U(0.42,0.45)$ & $N(75,2)$ &
      \citet{Orosz2003} \\
      GS 2023 & $N(6.08, 0.06)$ & $U(0.056,0.063)$ & $I(66, 70)$ &
      \citet{Charles2006} \\
      & & & & \citet{Khargharia2010} \\
      GS 1354 & $N(5.73, 0.29)$ & $N(0.12,0.04)$ & $I(50, 80)$ & 
      \citet{Casares2009} \\
      Nova Oph 77 & $N(4.86,0.13)$ & $U(0, 0.053)$ & $I(60, 80)$ &
      \citet{Charles2006} \\
      GS 2000 & $N(5.01, 0.12)$ & $U(0.035, 0.053)$ & $I(43, 74)$ &
      \citet{Charles2006} \\
      \hline \hline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% High Mass Follows
      Cyg X1 & $N(0.251, 0.007)$ & $N(2.778, 0.386)$ & $I(23, 38)$ &
      \citet{Gies2003} \\ 
      M33 X7 & $N(0.46, 0.08)$ & $N(4.47, 0.61)$ & $N(74.6, 1)$ &
      \citet{Orosz2007} \\
      NGC 300 X1 & $N(2.6, 0.3)$ & $U(1.05, 1.65)$ & $I(60, 75)$ &
      \citet{Crowther2010} \\
      LMC X1 & $N(0.148, 0.004)$ & $N(2.91, 0.49)$ & $N(36.38, 2.02)$
      & \citet{Orosz2009} \\
      IC 10 X1 & $N(7.64, 1.26)$ & $U(0.7, 1.7)$ & $I(75, 90)$ & 
      \citet{Prestwich2007} \\
       & & & & \citet{Silverman2008} \\
       \hline
    \end{tabular}
  \end{center}
  \caption{\label{tab:sources} The source parameters for the 20 low
    mass X-ray binaries used in this work.  The first 15 systems have
    low-mass secondaries that feed the black hole via Roche lobe
    overflow; the last five systems have high-mass secondaries ($q
    \gtrsim 1$) that feed the black hole via winds.  In each line, $f$
    is the mass function for
    the compact object, $q$ is the mass ratio $M_2/M$, and $i$ is the
    inclination of the system to the line of sight.  We indicate the
    distribution used for the true parameters when computing the
    probability distributions for the masses of these systems:
    $N(\mu,\sigma)$ implies a Gaussian with mean $\mu$ and standard
    deviation $\sigma$, $U(a,b)$ is a uniform distribution between $a$ and
    $b$, and $I(\alpha,\beta)$ is an isotropic distribution between the
    angles $\alpha$ and $\beta$.}
\end{table}

From these assumptions, we can generate probability distributions for
the true mass of the black hole given the observations and errors via
the Monte Carlo method: drawing samples of $f$, $q$, and $i$ from the
assumed distributions and computing the mass implied by Equation
\eqref{eq:mass-function} gives samples of $M$ from the distribution
induced by the relationship in Equation \eqref{eq:mass-function}.
Mass distributions generated in this way for the systems used in this
work are shown in Figure \ref{fig:low-masses} and Figure
\ref{fig:high-masses}.  Note that systems for which $i$ is poorly
constrained have broad ``tails'' on their mass distributions.  These
mass distributions constitute the ``observational data'' we will use
in the remainder of this paper.

@@

  \caption{\label{fig:low-masses} The mass distributions implied by
    Equation \eqref{eq:mass-function} and the assumed distributions on
    observational parameters given in Table \ref{tab:sources} for the
    lower-mass sources.  The significant asymmetry and long tails in
    many of these distributions are the result of the non-linear
    relationship (Equation \eqref{eq:mass-function}) between $M$, $f$,
    $q$, and $i$.}
\end{figure}

\begin{figure}
  \begin{center}
    \plotone{plots/high-masses}
  \end{center}
  \caption{\label{fig:high-masses} Mass distributions for the
    wind-fed, higher-mass systems computed from the distributions on
    observed data in Table \ref{tab:sources} using Equation
    \eqref{eq:mass-function}.  The asymmetry and long tails in these
    distributions are the result of the non-linear relationship
    between $M$, $f$, $q$, and $i$.}
\end{figure}

In Figure \ref{fig:masses} we show the combination of these individual
mass distributions into one single distribution.  Note that this
combination is not a good estimator of the true mass distribution of
black hole masses.  To understand why this is not a good estimator of
the underlying mass distribution, suppose that the observations
implied an identical mass distribution for each object.  Then the
combined distribution is equal to the individual distributions, but in
fact we can be confident that each of the objects was drawn from a
distribution that was narrower by a factor of approximately
$\sqrt{N}$.  \citet{Mandel2010} discusses this issue in detail.  We
include Figure \ref{fig:masses} here only to give a sense of the range
of masses of black holes in our sample; we will come to estimators for
the mass distribution.

\begin{figure}
  \plotone{plots/masses-high}

  \caption{\label{fig:masses} The mass distributions in Figures
    \ref{fig:low-masses} and \ref{fig:high-masses} combined into a
    single distribution. Note that this combination is not a good
    estimator of the true mass distribution of black hole masses in
    the galaxy---see, e.g., \citet{Mandel2010}.  We include it here
    only to give a sense of the range and general shape of the mass
    distribution of black holes in our sample.}
\end{figure}

@@

In this section we apply a Bayesian analysis to various models for the
underlying mass distribution from which the low-mass systems in Table
\ref{tab:sources} were drawn.  We will treat the combined low- and
higher-mass systems in Section \ref{sec:higher-mass}.  The end result
will be the probability distribution for the parameters of each model
implied by the data from Section \ref{sec:systems} in combination with
our prior assumptions about the probability distribution for the
parameters.  Bayes' rule relates these quantities.  For a model with
parameters $\vtheta$ in the presence of data $d$, Bayes' rule states
\begin{equation}
  \label{eq:Bayes-rule}
  p(\vtheta | d) = \frac{p(d | \vtheta) p(\vtheta)}{p(d)}.
\end{equation}
Here, $p(\vtheta|d)$, called the posterior probability distribution
function, is the probability distribution for the parameters $\vtheta$
implied by the data $d$; $p(d|\vtheta)$, called the likelihood, is the
probability of observing data $d$ given that the model parameters are
$\vtheta$; $p(\vtheta)$, called the prior, reflects our estimate of the
probability of the various model parameters in the absence of any
data; and $p(d)$, called the evidence, is an overall normalizing
constant ensuring that
\begin{equation}
  \int d\theta\, p(\vtheta|d) = 1,
\end{equation}
whence
\begin{equation}
  \label{eq:evidence-def}
  p(d) = \int d\vtheta\, p(d|\vtheta) p(\vtheta).
\end{equation}

In our context, the data are the mass distributions given in Section
\ref{sec:systems}: $d = \{ p_i(M)| i = 1, 2, \ldots, 17 \}$.  We
assume that the measurements in Section \ref{sec:systems} are
independent, so the complete likelihood is given by a product of the
likelihoods for the individual measurements.  For a model with
parameters $\vtheta$ that predicts a mass distribution $p(M|\vtheta)$
for black holes, we have
\begin{equation}
  \label{eq:likelihood-def}
  p(d|\vtheta) = \prod_i \int dM\, p_i(M) p(M|\vtheta).
\end{equation}
That is, the likelihood of an observation is the average over the
individual mass distribution implied by the observation, $p_i(M)$, of
the probability for a black hole of that mass to exist according to
the model of the mass distribution, $p(M | \vtheta)$.  We approximate
the integrals as averages of $p(M|\vtheta)$ over the Monte Carlo mass
samples drawn from the distributions in Table \ref{tab:sources} (also
see Figures \ref{fig:low-masses} and \ref{fig:high-masses}):
\begin{equation}
  p(d|\vtheta) \approx \prod_i \frac{1}{N_i} \sum_{j = 1}^{N_i} p(M_{ij} | \vtheta),
\end{equation}
where $M_{ij}$ is the $j$th sample (out of a total $N_i$) from the
$i$th individual mass distribution.

Our calculation of the likelihood of each observation does not include
any attempt to account for selection effects in the observations.  We
simply assume (almost certainly incorrectly) that any black hole drawn
from the underlying mass distribution is equally likely to be
observed, leaving the analysis of selection effects to future work.
(During the preparation of this work, we became aware of
\citet{Ozel2010}, which finds that selection effects do not greatly
affect an analysis similar to ours.)

For a mass distribution with several parameters, $p(\vtheta | d)$
lives in a multi-dimensional space.  Exploring the entirety of this
space systematically rapidly becomes prohibitive as the number of
parameters increases.  A more efficient way to explore the
distribution $p(\vtheta | d)$ is to use a Markov Chain Monte Carlo
(MCMC) method.  MCMC methods produce a chain (sequence) of parameter
samples, $\{ \vtheta_i, i = 1, \ldots \}$, such that a particular
parameter set, $\vtheta$, appears in the sequence with a frequency
equal to its probability according to $p(\vtheta|d)$.  In this way,
regions of parameter space where $p(\vtheta|d)$ is large are sampled
densely while regions where $p(\vtheta|d)$ is small are effectively
ignored.  A Markov chain has the property that the transition
probability from one element to the next, $p(\vtheta_i \to
\vtheta_{i+1})$, depends only on the value of $\vtheta_i$, not on any
previous values in the chain.

One way to produce a sequence of MCMC samples is via the following
algorithm, first proposed by \citet{Metropolis1953}:
\begin{enumerate}
  \item Begin with the current sample, $\vtheta_i$.
  \item Propose a new sample, $\vtheta_p$, by drawing randomly from a
    ``jump proposal distribution'' with probability $Q(\vtheta_i \to
    \vtheta_p)$.  Note that $Q(\vtheta_i \to \vtheta_p)$ can depend on
    the current parameters, $\vtheta_i$, and any other ``constant''
    data, but cannot examine the history of the chain beyond the most
    recent point.  This is necessary to preserve the Markovian
    property of the chain.
  \item Compute the ``acceptance'' probability,
    \begin{equation}
      \label{eq:paccept}
      p_{\textnormal{accept}} \equiv
      \frac{p(\vtheta_p|d)}{p(\vtheta_i|d)} \frac{Q(\vtheta_p \to
        \vtheta_i)}{Q(\vtheta_i \to \vtheta_p)}
    \end{equation}
  \item With probability $\min(1,p_{\textnormal{accept}})$ ``accept''
    the propsed $\vtheta_p$, setting $\vtheta_{i+1} = \vtheta_p$;
    otherwise set $\vtheta_{i+1} = \vtheta_i$.
\end{enumerate}
This algorithm is more likely to accept a proposed jump when it
increases the posterior (the first factor in Equation
\eqref{eq:paccept}) and when it is to a location in parameter space
from which it is easy to return (the second factor in Equation
\eqref{eq:paccept}); the combination of these influences in Equation
\eqref{eq:paccept} ensures that the equilibrium distribution of the
chain is $p(\vtheta|d)$.  As $i \to \infty$ the samples $\vtheta_i$
are distributed according to $p(\vtheta|d)$.

In practice the number of samples required before the chain
appropriately samples $p(\vtheta|d)$ depends strongly on the jump
proposal distribution; proposal distributions that often propose jumps
toward or within regions of large $p(\vtheta|d)$ can be very
efficient, while poor proposal distributions can require prohibitively
large numbers of samples before convergence.

There is no foolproof test for the convergence of a chain.  In this
work, we test the convergence of our chains by comparing the
statistics calculated from the entire chain to statistics calculated
from only the first half of the chain; when the chain has converged,
the two calculations agree.  This is a necessary, but not sufficient,
condition for convergence.

We begin the chain at an arbitrary point in parameter space; this is
equivalent to taking a finite section of an infinite chain that begins
with the chosen point.  Since every point in parameter space occurs in
an infinite chain, and no section of the chain is better than any
other, our chosen finite chain will have the appropriate distribution.
However, because consecutive samples in a chain are correlated with
each other, the beginning of our finite chain has a ``memory'' of the
starting point; we discard enough points at the beginning of the
finite chain that we can be confident it does not retain a memory of
the arbitrary starting point.  The points discarded in this way are
commonly called ``burn-in'' points.

Once we have a chain of samples from $p(\vtheta|d)$---a distribution
of distributions---the distribution for any quantity of interest can
be computed.  For example, Figure \ref{fig:dists} shows the median
values, and the 10\% and 90\% quantiles for the value of the BH mass
probability distribution at various masses.  (Recall that we are only
using the 15 low-mass observations in this Section's analysis.)  Such
a figure can be generated from the MCMC output by computing the value
of the parameterized BH mass distribution at various masses for each
parameter sample output by the MCMC, and then plotting the quantiles
at each mass.

\begin{figure}
  \begin{center}
    \plottwo{plots/dist-parameteric}{plots/dist-non-parameteric}
  \end{center}
  \caption{\label{fig:dists} The median values of the black hole mass
    distribution, $p(M|\theta)$, at various masses implied by the
    posterior $p(\theta|d)$ for the models discussed in Sections
    \ref{subsec:parameteric-models} and
    \ref{subsec:non-parameteric-models}.  These distributions use only
    the 15 low-mass observations in Table \ref{tab:sources} (the
    complete sample is analyzed in Section \ref{sec:higher-mass}).
    Error bars span the 10\% to 90\% range.  Note that these
    ``distributions of distributions'' are not necessarily normalized,
    and need not be ``shaped'' like the underlying model
    distributions.}
\end{figure}

@@

\subsection{Priors}

An important part of any Bayesian analysis is the priors placed on the
parameters of the model.  The choice of priors can bias the results of
the analysis through both the shape and the range of prior support in
parameter space.
